---
title: DANL Project
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - Alban Kuka
  - Henry Socash
  - Victoria Pavlock
  - Jeremy Sun
  - Jackson Boyer 

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
earthquake <- read.csv('/Users/alban/Downloads/earthquake data.csv')
theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

About this project :clap: - My teammates and I will be observing worldwide earthquake occurrences data that took place in 2023. Analyzing this data will allow us to determine how Depth varies by Magnitude and how factors such as Location and time affect the destructiveness of these earthquakes.

```{r}
earthquake <- read.csv('/Users/alban/Downloads/earthquake data.csv')
```

## Significance

Analyzing this data will help us determine the countries across the world that encounter the most devastating earthquakes. This can help and raise awareness for natural disasters in these countries and contribute to safety protocols for when earthquakes occur. The more we can prove a country is suffering because of natural disasters the more we can advocate for advanced infrastructure designed for safety during these occurrences. 


# Data

The data.frame `mpg` contains a subset of the fuel economy data that the EPA makes available on [https://fueleconomy.gov/](https://fueleconomy.gov/). It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car. :oncoming_automobile:


## Summary Statistics

```{r}
earthquake <- read.csv('/Users/alban/Downloads/earthquake data.csv')
```

```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(earthquake) 
```


```{r}
earthquake1 <- earthquake %>% select(Magnitude,Depth)
skim(earthquake1)
```


## MPG and a Type of Cars
The following boxplot shows how the distribution of highway MPG (`hwy`) varies by a type of cars (`class`) :blue_car: :truck: :minibus:.


```{r}
ggplot(data = mpg) +
  geom_boxplot(aes(x = class, y = hwy, fill = class),
               show.legend = F) +
  labs(x = "Class", y = "Highway\nMPG") 
```

## people 
```{r}

```


